---
title: "brain_markdown"
author: "Ines Franzolin"
date: "2024-12-02"
output:
  word_document: default
  pdf_document: default
  html_document: default
---
loading of the data
```{r}
WD <- "/Users/inesfranzolin/Desktop/UNIGE/BLL/Master_thesis/data_brain_analysis/"
setwd(WD)
brain_analyses <- read.csv("final_brain.csv")
```

libraries
```{r}
library(pacman)
p_load(gamm4, Hmisc, corrplot, tidyr)
```

descriptive of the variables used 
```{r}
brain_df = data.frame(brain_analyses[, c("Age", "multilingual_score_comprehension", "multilingual_score_speaking", "Hand_index", "number_of_languages",
                                  "eTIV.x", "left_parstriangularis", "left_ATL", "lh_G_pariet_inf.Supramar_volume", 
                                  "lh_precuneus_volume", "rh_precuneus_volume", "Left.Hippocampus", "Right.Hippocampus",
                                  "lh_parahippocampal_volume", "rh_parahippocampal_volume", "lh_entorhinal_volume",
                                  "rh_entorhinal_volume", "Left.Caudate", "left_parsopercularis",
                                  "left_dorsolateralPFC", "left_SMA", "lh_caudalanteriorcingulate_volume",
                                  "rh_caudalanteriorcingulate_volume")])
brain_df$sex <- as.factor(brain_analyses$Sex)

sd(brain_df$Age)
sd(brain_df$number_of_languages)
summary(brain_df)
```

inhibition network brain area's volume
```{r}
#dorsolateral prefrontal cortex, volume from Brainnetome Atlas
brain_analyses$left_dorsolateralPFC <- rowSums(brain_analyses[, c("lh_A9l_L_volume", "lh_A9m_L_volume", "lh_46d_L_volume", 
                                                                  "lh_46v_L_volume", "lh_A46_L_volume")])
#SMA, volume from Brainnetome Atlas
brain_analyses$left_SMA <- rowSums(brain_analyses[, c("lh_A6dl_L_volume", "lh_A6m_L_volume", "lh_A6vl_L_volume")])

#pars opercularis, volume from Brainnetome Atlas
brain_analyses$left_parsopercularis <- rowSums(brain_analyses[, c("lh_A44d_L_volume", "lh_A44op_L_volume", "lh_A44v_L_volume")])
```

scaling and data frame creation for the inhibition network
```{r}
#scaling inhibition variables
scaling_function <- function(x) {
  scale(x)}

to_scale_inh = brain_analyses[, c("Age", "multilingual_score_speaking", "Hand_index", 
                                  "eTIV.x", "Left.Caudate", "left_parsopercularis",
                                  "left_dorsolateralPFC", "left_SMA", "lh_caudalanteriorcingulate_volume",
                                  "rh_caudalanteriorcingulate_volume")]

scaled_inh <- scaling_function(to_scale_inh)

#data frame creation inhibition network
inhibition_df <- data.frame(scaled_inh)
colnames(inhibition_df)= c("age", "multilingual_score_speaking", "handedness", "volume_total", 
                                                      "left_caudate", "left_parsopercularis", "left_DLPFC", "left_SMA", 
                                                      "left_ACC", "right_ACC")


inhibition_df$sex <- as.factor(brain_analyses$Sex)
inhibition_df$participant_ID <- as.factor(brain_analyses$participant_id)
#convert to long format
inhibition_df_long <- pivot_longer(inhibition_df, cols = c("left_caudate", "left_parsopercularis", "left_DLPFC",
                                                           "left_SMA", "left_ACC", "right_ACC"),
                                     names_to = 'brain_area_name', values_to = 'brain_area_volume')

inhibition_df_long$brain_area_name <- as.factor(inhibition_df_long$brain_area_name)

brain_areas_inh <- levels(inhibition_df_long$brain_area_name)
```

GAMM inhibition network
```{r}
inhibition_gam <- gamm4(brain_area_volume ~ s(multilingual_score_speaking, by= brain_area_name) + volume_total + handedness + age + sex, 
      random = ~(1|participant_ID), data = inhibition_df_long, family = gaussian() ,REML = T) 

summary_inhibition <- summary(inhibition_gam$gam)
summary(inhibition_gam$gam)
```

Plots: relationship between multilingualism speaking and brain areas related to inhibition network
```{r}
for (i in 1:length(brain_areas_inh)) {
  plot(inhibition_gam$gam, select = i, residuals = TRUE, shade = TRUE,
       shade.col = 'deeppink3', ylab = paste(brain_areas_inh[i], "volume"), 
       xlab = "multilingual score speaking")
}
```


declarative network brain area's volume
```{r}
#pars triangularis, volume from Brainnetome Atlas
brain_analyses$left_parstriangularis <- rowSums(brain_analyses[, c("lh_A45c_L_volume", "lh_A45r_L_volume")])

#Anterior temporal lobe, volume from Brainnetome Atlas
brain_analyses$left_ATL <- rowSums(brain_analyses[, c("lh_A38m_L_volume", "lh_A38l_L_volume", "lh_A21r_L_volume", "lh_A20r_L_volume")])
```


scaling and data frame creation for the declarative memory network
```{r}
to_scale_dec = brain_analyses[, c("Age", "multilingual_score_comprehension", "Hand_index", 
                                  "eTIV.x", "left_parstriangularis", "left_ATL", "lh_G_pariet_inf.Supramar_volume", 
                                  "lh_precuneus_volume", "rh_precuneus_volume", "Left.Hippocampus", "Right.Hippocampus",
                                  "lh_parahippocampal_volume", "rh_parahippocampal_volume", "lh_entorhinal_volume",
                                  "rh_entorhinal_volume")]

scaled_dec <- scaling_function(to_scale_dec)

#data frame creation declarative
declarative_df <- data.frame(scaled_dec)
colnames(declarative_df) = c("age", "multilingual_score_comprehension", "handedness", "volume_total", 
                             "left_parstriangularis", "left_ATL", "left_supramarginal", "left_precuneus",
                             "right_precuneus", "left_hippocampus", "right_hippocampus", "left_parahippocampus",
                             "right_parahippocampus", "left_enthorinal", "right_enthorinal")

declarative_df$sex <- as.factor(brain_analyses$Sex)
declarative_df$participant_ID <- as.factor(brain_analyses$participant_id)

#convert to long format 
declarative_df_long <- pivot_longer(declarative_df, cols = c("left_parstriangularis", "left_ATL", "left_supramarginal", "left_precuneus",
                                                             "right_precuneus", "left_hippocampus", "right_hippocampus", "left_parahippocampus",
                                                             "right_parahippocampus", "left_enthorinal", "right_enthorinal"), names_to = 'brain_area_name', values_to = 'brain_area_volume')

declarative_df_long$brain_area_name <- as.factor(declarative_df_long$brain_area_name)

brain_areas_dec <- levels(declarative_df_long$brain_area_name)
```

GAMM declarative memory network
```{r}
declarative_gam <- gamm4(brain_area_volume ~ s(multilingual_score_comprehension, by= brain_area_name) + volume_total + handedness + age + sex, 
                         random = ~(1|participant_ID), data = declarative_df_long, family = gaussian() ,REML = T)

summary_declarative <- summary(declarative_gam$gam)
summary(declarative_gam$gam)
```

Plots: relationship between multilingualism comprehension and brain areas related to declarative memory network
```{r}
for (i in 1:length(brain_areas_dec)) {
  plot(declarative_gam$gam, select = i, residuals = TRUE, shade = TRUE,
       shade.col = 'deeppink3', ylab = paste(brain_areas_dec[i], "volume"), 
       xlab = "multilingual score comprehension")
}
```

Correlations between areas from inhibition and declarative memory network
```{r}
#dataframe creation
corr_inhdec <- data.frame(brain_analyses$lh_precuneus_volume, brain_analyses$rh_precuneus_volume,
                          brain_analyses$Left.Hippocampus,brain_analyses$Right.Hippocampus,
                          brain_analyses$lh_entorhinal_volume, brain_analyses$rh_entorhinal_volume,
                          brain_analyses$lh_parahippocampal_volume, brain_analyses$rh_parahippocampal_volume,
                          brain_analyses$left_parstriangularis,
                          brain_analyses$left_ATL,
                          brain_analyses$lh_G_pariet_inf.Supramar_volume,
                          brain_analyses$Left.Caudate, 
                          brain_analyses$left_parsopercularis,
                          brain_analyses$left_dorsolateralPFC, 
                          brain_analyses$lh_caudalanteriorcingulate_volume,
                          brain_analyses$rh_caudalanteriorcingulate_volume,
                          brain_analyses$left_SMA)

names(corr_inhdec) <- c(  "L PCun", "R PCun", "L HI", "R HI",
  "L Ent", "R Ent", "L PHG", "R PHG",
  "L IFGTr", "L ATL", "L SMG", "L Cd",
  "L IFGOp", "L DLPFC", "L ACC", "R ACC", "L SMA")
```

```{r}
#correlation matrice
corr_inhdec <- as.matrix(corr_inhdec)

matrice_inhdec <- rcorr(corr_inhdec, type = c('spearman'))
rho_inhdec = matrice_inhdec$r 
p_values_inhdec = matrice_inhdec$P 

#Benjamin Hochsberg correction
p_values_inhdec_vector <- as.vector(p_values_inhdec)
corrected_p_inhdec_vector = p.adjust(p_values_inhdec_vector, method = c('BH'), n = length(p_values_inhdec_vector))
corrected_p_inhdec <- matrix(corrected_p_inhdec_vector, nrow = nrow(p_values_inhdec),
                          dimnames = list(rownames(p_values_inhdec), colnames(p_values_inhdec))) 
diag(corrected_p_inhdec) <- 1
```

Plot: correlations between brain areas from inhibition and declarative network
```{r}
inhdec_corr_plot <- corrplot(rho_inhdec, method= 'circle', type = 'lower', order = 'original', 
                          p.mat = corrected_p_inhdec, sig.level = 0.05, insig = 'blank', 
                          addCoef.col = 'black', number.cex = 0.8, diag = F)
```


rho and p-values
```{r}
#rho
rho_inhdec

#p-values
corrected_p_inhdec
```



